Let $\mathrm{E} = (E,\#,\vdash)$ be an event structure where
$E = \s{e_1, e_2, ...,e_n}$ and $\mathcal{C}$ be a subset of
$\mathcal{P}(E)$, we define the causal model
$\mathfrak{M}(\mathrm{E},\mathcal{C}) = (\mathcal{S},\mathscr{F},\mathcal{E})$
where
$\mathcal{S} = (\mathcal{U},\mathcal{V},\mathcal{R})$.
We define $\mathcal{U}$ to be empty and $\mathcal{V}$ as follows:
\begin{align*}
    \mathcal{V} = & \s{C_{e_i,e_j} |  1 \leq i < j \leq n.
    e_i \in E \amp e_j \in E}                                \\
                  & \cup \s{EN_{s,e} | s \in \mathcal{P}(E),
    e \in E. e \not \in s }                                  \\
                  & \cup \s{M_{s,e} | s \in \mathcal{P}(E),
        e \in E. e \not \in s } \cup \s{ES,IC}
\end{align*}
Let $\mathcal{R}(ES) = \mathbb{E}$ where $\mathbb{E}$ is the set of all
triples of the form $(E',\#',\vdash')$ where $\#' \subseteq E \times E$
and $\vdash' \subseteq \mathcal{P}(E') \times E'$ and we define all other
variables to be boolean.
For each variable $x \in \mathcal{V}$ we define $\vec V_x$ as a vector
of all variables in $\mathcal{V}$ excluding $x$.
For $x,y \in \mathcal{P}(E)$ we say $x$ is covered by $y$ written $ x \prec y$ iff:
\begin{align*}
    x \subseteq y \amp x \neq y \amp
    (\forall z. x \subseteq z \subseteq y \Rightarrow x = z
    \text{ or } y = z)
\end{align*}
We define the functions in $\mathscr{F}$ as follows:
$$
    \f{C_{e,e'}} = \begin{cases}
        true  & \text{ if } e \# e' \amp e' \# e \\
        false & \text{ otherwise }
    \end{cases}
$$
$$
    \f{M_{s,e}} = \begin{cases}
        Min(s,e) \wedge Con(s) & \text{ if } s \vdash_{min} e \\
        false                  & \text{ otherwise }
    \end{cases}
$$
\begin{align*}
    \f{EN_{s,e}} & =
    \left(
    M_{s,e} \vee
    \left(
    \bigvee_{s'\prec s}EN_{s',e}
    \right)
    \right)
    \bigwedge
    Con(s)
\end{align*}
Where we have:
\begin{align*}
    Con(s)   & =   \left(
    \bigwedge_{ 1\leq j<j' \leq n \wedge e_j,e_{j'} \in s}
    \neg C_{e_j,e_{j'}}
    \right)               \\
    Min(s,e) & = \left(
    \bigwedge_{s'. (s' \subset s \vee s \subset s')
        \wedge e \notin s'}
    \neg M(k',i)
    \right)
\end{align*}
Additionally, we define:
\begin{align*}
    \f{ES} = (E,\#',\vdash')
\end{align*}
where we have:
\begin{align*}
    \forall e,e' \in E. e \#' e' \wedge e' \#' e
     & \iff C_{e,e'} = \T \\
    \forall s \in \mathcal{P}(E), e \in E.  s \vdash' e
     & \iff EN_{s,e} = \T
\end{align*}
Intuitively, the value of $ES$ represents the event structure induced
regarding the current setting of variables.
We define the function of $IC$ to determine whether $ES$
contains any element of $\mathcal{C}$:
\begin{align*}
    \f{IC} = \bigvee_{c \in \mathcal{C}}c \in \mathcal{F}(ES)
\end{align*}
Note that since we have defined $\mathcal{U}$ to be empty and
\textcolor{red}{the model is recursive}, the values of all variables in
$\mathcal{V}$ can be uniquely determined in the model.

We define $\mathcal{E}$ to be the set of all allowable
settings of the endogenous variables such as $\vec v'$
for which if $M \vDash \vec V = \vec v' \wedge ES = \mathrm{E}'$
then $\mathrm{E}'$ is an event structure.

\subsection{Actual Cause of Property Violation}
\begin{definition}
    Given a set of events $E$, we define a safety property $P$ to be a
    subset of $\mathcal{P}(E)$.
\end{definition}
\begin{definition}
    Given an event structure $\mathrm{E}$ we say it is safe regarding the 
    property $P$ if $\mathcal{F}(\mathrm{E}) \subseteq P$.
\end{definition}
Let $E$ be set of events and $\mathrm{E} = (E,\#,\vdash)$ be an event structure.
Let $P \subseteq \mathcal{P}(E)$ be a safety property that is violated in 
$\mathrm{E}$.
The property is violated if any $c \in \mathcal{P}(E) - P$ be a configuration
of $\mathrm{E}$. 
So, to find the actual cause of the property violation we need to find 
the actual cause of 
$\bigvee_{c \in \mathcal{P}(E) - P}c \in \mathcal{F}(\mathrm{E})$.
\begin{definition}
    Given a set of events $E$, let $P$ be a safety property on $E$ and 
    $\mathrm{E} = (E,\#,\vdash)$ be an event structure where $P$ is violated.
    If $\mathcal{M} = \mathfrak{M}(\mathrm{E}, P - \mathcal{P}(E))$
    be a causal model then $\vec X = \vec x$ is
    an actual cause of the violation of $P$ in $\mathrm{E}$ if the
    following three conditions hold:
    \begin{itemize}
        \item  \textbf{AC1.} $M\models \vec X = \vec x
                  \wedge \bigvee_{c \in \mathcal{P}(E) - P}c \in \mathcal{F}(\mathrm{E})$.
        \item  \textbf{AC2. }There exists a partition $(\vec Z, \vec W)$ of $\mathcal{V}$ with $\vec X \subseteq \vec Z$ and some setting $(\vec x',\vec w')$ of the variables in $(\vec X,\vec W)$ such that if $(M,\vec u)\models \vec Z = z^*$ for all $Z\in \vec Z$, then both of the following conditions hold:

              (a) $M \models[\vec X \leftarrow \vec x', \vec W \leftarrow \vec w']
                  IC = \F
                  \wedge \vec V = \vec v
                  \wedge  \vec v \in \mathcal{E}$.

              (b) $M \models[\vec X\leftarrow \vec x, \vec W' \leftarrow \vec w', \vec Z'\leftarrow \vec z^*]
                  IC = \T
                  \wedge \vec V = \vec v
                  \wedge \vec v \in \mathcal{E}$
              for all subsets $\vec W'$ of $\vec W$ and all subsets $Z'$ of $\vec Z$.

        \item  \textbf{AC3.} $\vec X$ is minimal; no subset of $\vec X$ satisfies conditions $AC1$ and $AC2$.
    \end{itemize}
    Where $\vec v$ is the value of endogenous variables.
\end{definition}
\pagebreak


\subsection{Examples}
\input{sections/examples}

\subsection{Causality in DyNetKAT}
\input{sections/dynetkat.tex}