\begin{definition}[Restriction]
    Let $\mr{E} = (E,\#,\vdash)$ be an event structure.
    Let $A \subseteq E$.
    Define the restriction of $\mr{E}$ to $A$ to be
    \begin{align*}
        \mr{E} \lceil A = (A,\#_A,\vdash_A)
    \end{align*}
    where
    \begin{align*}
        X \in Con_A \iff X \subseteq A \ \wedge \ X \in Con \\
        X \vdash_A e \iff X \subseteq A \ \wedge \ e \in A \ \& \ X \vdash e
    \end{align*}
\end{definition}

\begin{definition}[Prefixing]
    Let $a$ be an event.
    For an event structure $\mr{E} = (E,\#,\vdash)$ define $a\mr{E}$ to be the event structure $(E',\#',\vdash')$ where:
    \begin{align*}
         & E' = \s{(0,a)} \cup \s{(1,e)|e \in E},                                                                                                        \\
         & \forall e_0',e_1'\in E' . e_0' \#' e_1'  \iff \exists e_0,e_1.e_0' = (1,e_0)
        \ \wedge \ e_1' = (1,e_1) \ \wedge \ e_0 \# e_1                                                                                                  \\
         & \forall X \subseteq E' . X \vdash' e' \iff e' = (0,a) \text{ or } [e' = (1,e_1) \ \wedge \ (0,a)\in X \ \wedge \ \s{e|(1,e)\in X} \vdash e_1]
    \end{align*}
\end{definition}

\subsubsection{Prefix}
\begin{definition}[Prefix]
    Let $(E,\#,\vdash,L,l)$ be a labelled event structure.
    Let $\alpha$ be a label.
    Define $\alpha(E,\#,\vdash,L,l)$ to be a labelled event structure
    $(E',\#',\vdash',L',l')$ where $(E',\#',\vdash') = \alpha(E,\#,\vdash)$
    and labels are defined as follows:
    \begin{align*}
        L' = \s{\alpha} \cup L
    \end{align*}
    and
    $$
        l'(e') = \begin{cases}
            \alpha & \text{if } e' = (0,\alpha) \\
            l(e)   & \text{if } e' = (1,e)
        \end{cases}
    $$
    for all $e' \in E'$.
\end{definition}

\subsubsection{Sum}
\begin{definition}[Sum]
    Let $\mr{E_0} = (E_0,\#_0,\vdash_0,L_0,l_0)$ and
    $\mr{E_1} = (E_1,\#_1,\vdash_1,L_1,l_1)$ be labelled event structures.
    Their sum $\mr{E_0} + \mr{E_1}$, is defined to be the structure $(E,\#,\vdash,l)$
    with events $E = \s{(0,e)|e \in E_0} \cup \s{(1,e)|e \in E_1}$,
    the disjoint union of sets $E_0$ and $E_1$,
    with injections $\iota_k: E_k \rightarrow E$, given by
    $\iota_k(e) = (k,e)$, for $k=0,1$, conflict relation
    \begin{align*}
        \forall e,e' \in E.
        e \# e' \iff & \exists e_0,e_0' \in E_0. e = \iota_0(e_0)
        \wedge e' = \iota_0(e_0') \wedge e_0 \#_0e_0'                                                      \\
                     & \text{or } \exists e_1,e_1' \in E_1. e = \iota_1(e_1) \wedge
        e' = \iota_1(e_1') \wedge e_1 \#_1 e_1'                                                            \\
                     & \text{or } \exists e_0 \in E_0,e_1 \in E_1.(e=\iota_1(e_0) \wedge e' =\iota_1(e_1)) \\
                     & \text{or } (e'=\iota_1(e_0) \wedge e =\iota_1(e_1))
    \end{align*}
    and enabling relation
    \begin{align*}
        X \vdash e \iff & X \in Con \wedge e \in E \wedge                   & \\
                        & (\exists X_0 \in Con_0,e_0 \in E_0.X = \iota_0X_0
        \wedge e = \iota_0(e_0) \wedge X_0 \vdash_0 e_0) \text{ or }          \\
                        & (\exists X_1 \in Con_1,e_1 \in E_1.X = \iota_1X_1
        \wedge e = \iota_1(e_1) \wedge X_1 \vdash_1 e_1)                      \\
    \end{align*}
    Where $Con_0,Con_1$ are defined on $\mr{E}_0$ and $\mr{E}_1$ respectively.
    We define the set of labels as $L_0 \cup L_1$ and the labelling function as:
    $$
        l(e) = \begin{cases}
            l_0(e_0) & \text{ if } e = \iota_0(e_0) \\
            l_1(e_1) & \text{ if } e = \iota_1(e_1)
        \end{cases}
    $$
\end{definition}

\subsubsection{Product}
In the product of two event structures, their events of synchronization are those pairs of events $(e_0,e_1)$, one from each event structure;
if $e_0$ is labelled $\alpha_0$ and $e_1$ is labelled $\alpha_1$ the synchronization event is
then labelled $(\alpha_0,\alpha_1)$.
Events need not synchronize however; an event in one component may not synchronize with
any event in the other.
We shall use events of the form $(e_0,*)$/$(*,e_1)$ to stand for the occurrence of an event $e_0$/$e_1$
from one component unsynchronized with any event of the other.
Such an event will be labeled by $(\alpha_0,*)$ where $\alpha_0$ is the original label of $e_0$
and * is a sort of undefined.

\begin{definition}[Product]
    Let $\mr{E_0} = (E_0,\#_0,\vdash_0,L_0,l_0)$ and
    $\mr{E_1} = (E_1,\#_1,\vdash_1,L_1,l_1)$
    be labeled event structures.
    Define their product $\mr{E_0} \times \mr{E_1}$ to be the structure $\mr{E} = (E,\#,\vdash,L,l)$
    consisting of events $E$ of the form
    \begin{align*}
        E_0 \times_* E_1 =
        \s{(e_0,*)|e_0 \in E_0}
        \cup \s{(*,e_1)|e_1 \in E_1}
        \cup \s{(e_0,e_1)| e_0 \in E_0 \wedge e_1 \in E_1}
    \end{align*}
    with projections $\pi_i : E \rightarrow_* E_i$,
    given by $\pi_i(e_0,e_1) = e_i$, for $i=0,1$, reflexive conflict relation $\doublevee$ given by
    \begin{align*}
        e \doublevee e' \iff \pi_0(e) \doublevee_0 \pi_0(e') \text{ or }
        \pi_1(e) \doublevee_1 \pi_1(e')
    \end{align*}
    for all $e,e' \in E$ we use $Con$ for the conflict-free finite sets,
    enabling relation $\vdash$ given by
    \begin{align*}
         & X \vdash e \iff X \in Con \wedge e \in E \wedge                  \\
         & (\pi_0(e)\text{ is defined } \Rightarrow \pi_0X\vdash_0\pi_0(e))
        \wedge (\pi_1(e)\text{ is defined } \Rightarrow \pi_1X\vdash_1\pi_1(e))
    \end{align*}
    Its set of labels is
    \begin{align*}
        L_0 \times_* L_1 = \s{ (\alpha_0,*)|\alpha_0 \in L_0}
        \cup \s{(*,\alpha_1)|\alpha_1 \in L_1}
        \cup \s{(\alpha_0,\alpha_1)|\alpha_0 \in L_0 \wedge \alpha_1 \in L_1}
    \end{align*}
    with projections: $\lambda_i: E \rightarrow_* E_i$ given by
    $\lambda_i(\alpha_0,\alpha_1) = \alpha_i$, for $i=0,1$.
    Its labeling function is defined to act on an event $e$ so
    \begin{align*}
        l(e) = (l_0\pi_0(e),l_1\pi_1(e))
    \end{align*}
\end{definition}

\subsubsection{Restriction}

\begin{definition}[Restriction]
    Let $\mr{E} = (E,\#,\vdash,L,l)$ be a labelled event structure.
    Let $\Lambda$ be a subset of labels.
    Define the restriction $\mr{E}\lceil \Lambda$ to be $(E',\#',\vdash',L\cap \Lambda,l')$
    where $(E',\#',\vdash')$ is the restriction of $(E,\#,\vdash)$
    to events $\s{e \in E|l(e) \in \Lambda}$ and the labeling function $l'$
    is the restriction of the original labeling function to the domain $L \cap \Lambda$.
\end{definition}

\section{Semantics of Communicating Processes}
In this section, first, we define a simple language
and we define the event structure semantics of this language.

% \subsection{Normal }

Let $\mc{A}$ be the set of possible labels for event structures.
We define a language with the following syntax:

\begin{align*}
    p ::= nil ~|~ \alpha p ~|~ p_0 + p_1 ~|~ p_0 \times p_1
    ~|~ p\lceil \Lambda
\end{align*}

where $\alpha \in \mc{A}$ and $\Lambda \subseteq \mc{A}$.

\subsection{Denotational Semantics}

\begin{definition}
    We define the semantic map $\sem{ \ }: p \ra \mathbb{E}$ where
    $\mathbb{E}$ is the set of all event structures with
    labels in $\mc{A}$ as follows:
    \begin{align*}
        \sem{nil}      & = (\emptyset,\emptyset)                  \\
        \sem{\alpha t} & = \alpha\sem{t}                        \\
        \sem{t_1 + t_2}
                        & = \sem{t_1} + \sem{t_2}                  \\
        \sem{t \lceil \Lambda}
                        & = \sem{t} \lceil \Lambda \\
        \sem{t_1 \times t_2}
                        & = \sem{t_1} \times \sem{t_2}
    \end{align*}
\end{definition}

\subsection{Examples}
In the following, we provide some examples to illustrate
how the prefix, sum, and product operators can be used to 
compose event structures.
For simplicity, assume that $ \mc{A} = \s{a,b,c}$. 

\begin{example}
    $\mr{E} = \sem{ab}$ is an event structure
    $(E,\#,\vdash,L,l)$ where:
    \begin{align*}
        E  & = \s{(0,a),(1,0,b)}                       \\
        \# & = \emptyset                               \\
           & \e \vdash (0,a), \s{(0,a)} \vdash (1,0,b) \\
        L  & = \s{a,b}                                 \\
           & l((0,a)) = a, l((1,0,b)) = b              
    \end{align*}
    The following diagram shows configurations of $\mr{E}$:
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            \crd[right]{0}{0}{$\emptyset$}
            \crd[right]{0}{1}{$\s{(0,a)}$}
            \crd[right]{0}{2}{$\s{(0,a),(1,0,b)}$}
            \draw [ultra thick] (0,0) -- (0,1);
            \draw [ultra thick] (0,1) -- (0,2);
        \end{tikzpicture}
    \end{center}
\end{example}

\begin{example}
    $\mr{E} = \sem{a + b}$ is an event structure
    $(E,\#,\vdash,L,l)$ where:
    \begin{align*}
        E & = \s{(0,a),(1,b)}                \\
          & (0,a) \# (1,b)   \\
          & \e \vdash (0,a), \e \vdash (1,b) \\
        L & = \s{a,b}                        \\
          & l((0,a)) = a, l((1,b)) = b       
    \end{align*}
    The following diagram shows configurations of $\mr{E}$:
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            \crd[above]{0}{0}{$\emptyset$}
            \crd[above]{-2}{1}{$\s{(0,a)}$}
            \crd[above]{2}{1}{$\s{(1,b)}$}
            \draw [ultra thick] (0,0) -- (-2,1);
            \draw [ultra thick] (0,0) -- (2,1);
        \end{tikzpicture}
    \end{center}
\end{example}


\begin{example}
    $\mr{E} = \sem{a \times b}$ is an event structure
    $(E,\#,\vdash,L,l)$ where:
    \begin{align*}
        E  & = \s{(a,*),(*,b),(a,b)}                           \\
        \# & = \e                                              \\
           & \e \vdash (a,*), \e \vdash (*,b),\e \vdash (a,b)  \\
        L  & = \s{(a,*),(*,b),(a,b)}                           \\
           & l((a,*)) = (a,*), l((*,b)) = (*,b),l((a,b))=(a,b) 
    \end{align*}
    The following diagram shows configurations of $\mr{E}$:
    \begin{center}
        \begin{tikzpicture}[scale=0.8]
            \crd[above]{0}{0}{$\emptyset$}
            \crd[left]{-2}{1}{$\s{(a,*)}$}
            \crd[right]{2}{1}{$\s{(*,b)}$}
            \crd[above]{0}{2}{$\s{(a,*),(*,b)}$}
            \crd[above]{2}{3}{$\s{(a,b)}$}
            \draw [ultra thick] (0,0) -- (-2,1);
            \draw [ultra thick] (0,0) -- (2,1);
            \draw [ultra thick] (2,1) -- (0,2);
            \draw [ultra thick] (-2,1) -- (0,2);
            \draw [ultra thick] (0,0) -- (2,3);
        \end{tikzpicture}
    \end{center}
\end{example}

In \cite{hp} there is no formal definition of the actual cause
in extended causal models.
So, here we provide a new definition of the actual cause
in the context of extended causal models.


\begin{definition}
    $\vec X = \vec x$ is an actual cause of $\varphi$ in $(M,\vec u)$ if the following three conditions hold:
    \begin{itemize}
        \item  \textbf{AC1.} $(M,\vec u)\models (\vec X = \vec x) \wedge \varphi \wedge $.
        \item  \textbf{AC2. }There exists a partition $(\vec Z, \vec W)$ of $\mathcal{V}$ with $\vec X \subseteq \vec Z$ and some setting $(\vec x',\vec w')$ of the variables in $(\vec X,\vec W)$ such that if $(M,\vec u)\models \vec Z = z^*$ for all $Z\in \vec Z$, then both of the following conditions hold:

              (a) $(M,\vec u)\models[\vec X \leftarrow \vec x', \vec W \leftarrow \vec w']\neg \varphi
                  \wedge \vec V = \vec v
                  \wedge  \vec v \in \mathcal{E}$.

              (b) $(M,\vec u)\models[\vec X\leftarrow \vec x, \vec W'
                      \leftarrow \vec w', \vec Z'\leftarrow \vec z^*]
                  \vec V = \vec v \wedge (\vec v \in \mc{E} \Rightarrow \varphi)$
              for all subsets $Z'$ of $\vec Z$.

        \item  \textbf{AC3.} $\vec X$ is minimal; no subset of $\vec X$ satisfies conditions $AC1$ and $AC2$.
    \end{itemize}
    Where $\vec v$ is the value of endogenous variables.
\end{definition}

Let $\mathbb{E}$ be the all triples of the form
$(E,\#',\vdash')$ on the set of events $E$ where
$\#' \subseteq E \times E$ and
$\vdash' \subseteq \mc{P}(E) \times E$.
We define a function
$ES: \times_{V \in \mc{V}\setminus \s{PV}} \mc{R}(V) \rightarrow \mathbb{E}$ which intuitively returns the event structure
that can be derived from the values of the endogenous variables
in the model excluding $PV$.
Let $\vec v$ be a vector of the values of the variables
in $\mc{V} \setminus \s{PV}$ and $\vec v(V)$ be the value of
$V$ in $\vec v$ for each $V \in \mc{V} \setminus \s{PV}$.
We define the function $ES$ so that if
$ES(\vec v) = (E,\#',\vdash')$ we have:
\begin{align*}
    \forall e,e' \in E. e \#' e' & \iff \vec{v}(C_{e,e'}) = \T \\
    \forall s \in \mathcal{P}(E), e \in E.  s \vdash' e
                                 & \iff \vec{v}(EN_{s,e}) = \T
\end{align*}
We define $\mathcal{E}$, the set of all allowable
settings of the endogenous variables, to be the vector of
values such as $\vec v'$ for which $ES(\vec v)$ is an
event structure.

Finally, to encode the property violation,
we define the function of $PV$ such that
if the value of the endogenous variables is
$\vec v$ then we have:
\begin{align*}
    PV = \T \iff S \cap \mc{F}(ES(\vec v)) = \e
\end{align*}

\subsection{Actual Cause of Unsafe Behavior}

Using the definition of the actual cause in the extended causal
model, we can rewrite the definition for unsafe behaviors in
event structures as follows:

\begin{definition}
    Let $\mr{E} = (E,\#,\vdash)$ be an event structure and
    $\mc{M}$ be the causal model of an unsafe behavior in
    $\mr{E}$ where unsafe behavior is specified as the
    function $F_{PV}$ in $\mc{M}$.
    We say $\vec X = \vec x$ is an actual cause of
    the unsafe behavior in $\mr{E}$ if the following
    conditions hold:
    \begin{itemize}
        \item  \textbf{AC1.} $M\models \vec X = \vec x
                  \wedge PV = \T$
        \item  \textbf{AC2. }There exists a partition $(\vec Z, \vec W)$ of $\mathcal{V}$ with $\vec X \subseteq \vec Z$ and some setting $(\vec x',\vec w')$ of the variables in $(\vec X,\vec W)$ such that if $(M,\vec u)\models \vec Z = z^*$ for all $Z\in \vec Z$, then both of the following conditions hold:

              (a) $M \models[\vec X \leftarrow \vec x', \vec W \leftarrow \vec w']
                  PV = \F
                  \wedge \vec V = \vec v
                  \wedge  \vec v \in \mathcal{E}$.

              (b) $M \models[\vec X\leftarrow \vec x, \vec W' \leftarrow \vec w', \vec Z'\leftarrow \vec z^*]
                  \vec V = \vec v
                  \wedge
                  (\vec v \in \mathcal{E} \Rightarrow PV = \T)
              $
              for all subsets $Z'$ of $\vec Z$.

        \item  \textbf{AC3.} $\vec X$ is minimal; no subset of $\vec X$ satisfies conditions $AC1$ and $AC2$.
    \end{itemize}
    Where $\vec v$ is the value of endogenous variables.
\end{definition}

We define the following programs:
\begin{align*}
    N & = p_i q_i f_{ad} + q_i p_i f_{ad} \\
    C & = p_o \times q_o \\
    SDN & = (N \times C)\lceil \Lambda \\
    \Lambda & = \s{(p_i,p_o),(q_i,q_o),(f_{ad},*)}
\end{align*}
This program captures the update behavior of the network 
figure \ref{fig:blacklist}.
Assume that two updates, $p$ and $q$ happening in the network.
$p$ replaces the path $ab$ with $ac$ and $q$ replaces
the path $cb$ with $cd$.
Let $C$ be the controller that concurrently sends commands for these
updates, $p_o,q_o$.
We use $p_i$ and $q_i$ for actions of receiving these commands and
$f_{ad}$ for the action of forwarding a packet from $a$ to $d$.
For simplicity, we do not consider any forwarding in the network 
other than $f_{ad}$.
Event structure of the $SDN$ has the configurations as depicted 
in figure \ref{fig:blacklist:es}.
Blacklisted nodes are nodes in the network that must not
be reachable \cite{network-abstractions}.
Let's assume the network in figure \ref{fig:blacklist} as
an example where the node $d$ is blacklisted.
For simplicity we assume that we require
$d$ not reachable only from $a$.
Thus, we consider the existence of configurations
$\s{p_1,q_1,ad_1}$ and $\s{p_2,q_2,ad_2}$ in an event structure
as a property violation.
In this example we can prove that $C_{p_1,q_1} = \F$ is an 
actual cause of the property violation.

\begin{example}
    Suppose that we want to reason about a forest fire that could
    be caused by either lightning or a match lit by an arsonist.
    Then the causal model would have the following endogenous variables :
    \begin{itemize}
        \item $F$ for fire
        \item $L$ for lighting
        \item $ML$ for match lit
    \end{itemize}
    The set $\mathcal{U}$ of exogenous variables includes conditions
    that suffice to render all relationships deterministic (such as
    whether the wood is dry, whether there is enough oxygen in the air
    for the match to light, etc.).
    Suppose that $\vec u$ is a setting of the exogenous variables that
    makes a forest fire possible (i.e., the wood is sufficiently dry,
    there is oxygen in the air, and so on).
    Then, for example, $F_F(\vec u, L, ML)$ is such that $F=1$ if either
    $L=1$ or $ML=1$.
    Note that although the value of $F$ depends on the value $L$ and $ML$, the value of $L$ does not depend on the values of $F$ and $ML$.
\end{example}